\input{111pre}
\begin{document}
\MYTITLE{Practical 5 \\ February 24, 2017\\Due in Bitbucket by midnight of the day of your practical \\ ``Checkmark'' grade}

\vspace*{-.2in}
\subsection*{Summary}
\vspace*{-.05in}

As a means of practicing with user input and output, the declaration of variables, and the creation of expressions in
the Java programming language, you will write a ``Mad Libs'' program involving words, numbers, and some simple
calculations. Then, using the {\tt git add}, {\tt git commit}, and {\tt git push} commands you should upload it to your
Git repository hosted by Bitbucket.

\noindent
If you have never heard of Mad Libs, then please visit this Web site: \url{http://www.madlibs.com/}.

\vspace*{-.15in}
\subsection*{Review the Textbook}
\vspace*{-.05in}

Be sure to read Sections 3.1 through 3.6 of your book to learn more about variables, data types, expressions, and user
input. As you review this material, try to make a list of questions about concepts that you do not yet fully
understand. In addition to discussing these questions with the teaching assistants and the course instructor, try to
answer them as you complete this assignment.

\vspace*{-.15in}
\subsection*{Exercise: Implement a ``Mad Libs'' Program}
\vspace*{-.05in}
Your program should ask the user to enter words and numbers, then print out a story using those words and numbers. The
numbers should be used to calculate something which will also be printed out. See Figure \ref{mad} for an example---but,
of course, you should create your own (much better) story!

\begin{figure}[tb]
\begin{Verbatim}[commandchars=\\\{\}]
aldenv27:gkapfham$ \textcolor{red}{javac MadLib.java}
aldenv27:gkapfham$ \textcolor{red}{java MadLib}
Gregory M. Kapfhammer, Practical 5
Thu March 3 13:11:17 EST 2016

Enter a singular noun: \textcolor{red}{noggin}
Enter an adjective: \textcolor{red}{verboten}
Enter another adjective: \textcolor{red}{glitzy}
Enter a non-zero whole number: \textcolor{red}{32}
Enter another non-zero whole number: \textcolor{red}{42}
Enter any number: \textcolor{red}{5.43}
Enter a singular verb: \textcolor{red}{snooze}

----------------------------------

Third-Grade Word Problem

If you own 32 verboten noggins,
and you purchase 42 glitzy noggins,
how many more noggins do you need to snooze 5.43 noggins?

Answer: You need -68.57 more noggins.
\end{Verbatim}
\vspace*{-.1in}
\caption{Sample ``Mad Libs'' output with user input highlighted in red.}
\label{mad}
\end{figure}

\noindent Remember to ``{\tt import java.util.Scanner}'' at the top of your Java program and to declare a {\tt Scanner}
variable (named, for instance, {\tt scan}---but you can name it something else if you want). Please use statements like
``{\tt... = scan.next()}'', ``{\tt... = scan.nextInt()}'', and ``{\tt... = scan.\\nextDouble()}'' to read in strings,
integers, and double values. Your program should produce output that is neat and your program's source code must make
good use of white space and labeling.

\vspace*{-.15in}
\subsection*{General Guidelines for Practical Sessions}
\vspace*{-.05in}
\begin{itemize}

\item {\bf Experiment!} Practical sessions are for learning by doing without the pressure of grades or ``right/wrong''
  answers. So try things!  The best way to learn is by trying things out.

\item {\bf Practice Key Laboratory Skills.} As you are completing this assignment, practice using the {\tt gvim} text
  editor and the Ubuntu terminal until you can easily run their most important features.  Additionally, ask
  a teaching assistant or the course instructor to teach you some of the advanced features of {\tt gvim} and the
  terminal, thereby helping you to work more effectively.

\item {\bf Try to Finish During the Class Session.} Practical exercises are not intended to be the equal of the
  laboratory assignments. If you are simply a slow typist, I've given you until the end of the day, but ideally you
  should upload a file, even a non-working one, by the end of the class period. You also should ensure that, for this
  assignment and all subsequent assignments, you can confidently upload files to your Git repository during the
  practical session.

\item {\bf Help One Another!} If your neighbor is struggling and you know what to do, offer your help. Don't ``do the
  work'' for them, but advise them on what to type or how to handle things. If you are stuck on a part of this practical
  session and you could not find any insights in either your textbook or online sources, formulate an intelligent
  question to ask your neighbor, a teaching assistant, or the course instructor. Try to strike the right balance between
  asking for help when you cannot solve a problem and working independently to find a solution.

\item {\bf Update Your Repository Often!} You should {\tt add}, {\tt commit}, and {\tt push} your updated files each
  time you work on them, always including descriptive messages about each code change.

\item {\bf Review the Honor Code Policy on the Syllabus.} Remember that while you may discuss programs with other
  students in the course, programs that are nearly identical to, or merely variations on, the work of others will be
  taken as evidence of violating the Honor Code.

\item {\bf Add an Example of Your ``Mad Lib'' to Slack.} Once you have finished your program and run it in the terminal
  window, then please paste your output into the \#practical channel on our Slack team. This will allow your colleagues
  to see how you have creatively used variables and expressions to create a MadLib system in the Java programming
  language.

\end{itemize}
\end{document}
